Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)

### 2015-08-23 research_class/paper.tex
paper.tex: spell check.

\author{Joel D. Martin}
\author{David Levine}
\affil{University of Texas at Arglington}

@@

\begin{abstract}
We describe the creation of a browser-based distributed peer-to-peer
application using WebRTC, PeerJS and Raft.js. WebRTC is a protocol that modern
browsers have begun to integrate which enables browser-to-browser
communication. PeerJS is a WebRTC client library and extensible signalling
server. Raft.js is an implementation in JavaScript of the Raft distributed
consensus protocol. These technologies are combined to create a chat
application that guarantees a shared, consistent and linearized view of the
chat history without relying on a single centralized server. We also create a
series of automated tests using the SlimerJS browser automation tool and use
them to show the correctness of the chat application and also to measure the
performance and scalability of the application over various cluster sizes.
\end{abstract}

@@

The Raft algorithm breaks the problem of consensus into three core
subproblems: leader election, log replication, and safety. A full Raft
solution must also address: cluster membership changes, log
compaction, and client interaction. The Raft algorithm implemented and
discussed in this paper is based on "Consensus: Bridging Theory and
Practice" by Diego Ongaro (PhD Dissertation). % TODO/CITE

@@

The current Raft.js implementation is based on "Consensus", Ongaro and
implements most of the Raft algorithm including: leader election, log
replication, safety, membership changes, basic client interaction, and
read-only operation optimizations. Log compaction and full client
linearizability are not currently implemented (clients and client
transactions are not assigned unique IDs). Log compaction is important
for long-running production systems in order to limit memory growth
and allow new nodes to catch up their logs more efficiently. However
log compaction is not required for verifying overall system
correctness and does not significantly impact short-running test
configurations. Full client linearizability is important feature for
simplifying certain types of distributed applications, however the
implementation requires some significant additions to the Raft
implemenation (e.g. assigning client IDs and sequence IDs and tracking
timed-out clients). For the chat application, simply adding one atomic
check-and-set state machine command is sufficient to achieve
a linearized chat history. This does impose a slight additional cost
on the client because it must retry the atomic operations until they
succeed. The author intends to implement both log compaction and
full client linearizability in the future but, however these do not
have a significant impact on the ability to test Raft combined with
WebRTC.

@@

\item Dynamic Raft.js over WebRTC: enhance the signaling server with support for multiple sessions/channels to enable separate Raft clusters with the same signalling server. Start each cluster with a single node and automatically add and remove Raft nodes as they come and go.

@@

In the Raft over WebRTC implemenation each local browser is a client
of the Raft cluster. However in Raft, the leader node is the only node
that adds entries to the log so clients on follower nodes must have
a way to communicate with the leader. This is implemented as pair of
functions named clientRequest and clientRequestResponse. clientRequest
is called by the client with the arguments of a client request RPC and
a callback function to be called once the request is completed. The
clientRequest/clientRequestResponse functions work as follows:

@@

\begin{enumerate}
    \item \emph{test/common.js}: a module of common functions used by
        the other tests scripts.
    \item \emph{test/test\_up.js}: launch and start an initial leader
        node/page, then start the rest of the nodes (if any) so that
        the leader will configure them into the cluster. Then test
        then waits for every node to be recorded (applied) into the log
        of every other node of the cluster. The test reports the
        number of milliseconds for the cluster to reach this state.
    \item \emph{test/test\_kill\_nodes.js}: similar to
        \emph{test/test\_up.js} but once all the nodes are configured
        into the cluster, this script then kills the leader plus
        a configurable number of nodes. By default the script kills
        just under half the nodes of the cluster. The test then waits
        for a new leader to be elected and remove the dead nodes from
        the cluster. The test reports both the number of milliseconds
        for the cluster to come up and the recover from the failure.
    \item \emph{test/test\_chat\_propagate.js}: similar to
        \emph{test/test\_up.js} but once the cluster is up, the test
        connects to one of the non-leader nodes and and sends a chat
        text line. The test then waits for the chat message to
        propogate to every other node's state machine. The test
        reports both the number of milliseconds for the cluster to
        come up and for the messages to propagate to every node.
\end{enumerate}

@@

\begin{enumerate}
    \item \emph{CPUs}: 8 Intel Core i7 processors running at 3.20GHz
    \item \emph{Memory}: 12 GB
    \item \emph{OS}: Ubuntu 14.04.1 LTS
    \item \emph{Docker}: version 1.3.1
    \item \emph{Docker image}:
    \begin{enumerate}
        \item \emph{OS}: Ubuntu 12.04.5 LTS
        \item \emph{SlimerJS}: vesion 0.9.6
        \item \emph{XULRunner}: 38.0.5
    \end{enumerate}
\end{enumerate}

@@

\section*{Acknowledgements}
I would like to thank my academic supervisor, David Levine, for his
excellent input and oversight of this project. I would also like to
thank Diego Ongaro for his amazing work on the Raft protocol and paper
which made implementing a Raft implementation a pleasure. There are
too many people involved in the creation and standardization of WebRTC
but I would like to thank them for their tireless work to make
browser-to-browser communication possible. Last but not least, I would
like to thank my family for putting up with the long nights and
weekends I spent on this project. Thanks!

>>>>

\author{Joel D. Martin}
\author{David Levine}
\affil{University of Texas at Arlington}

@@

\begin{abstract}
We describe the creation of a browser-based distributed peer-to-peer
application using WebRTC, PeerJS and Raft.js. WebRTC is a protocol that modern
browsers have begun to integrate which enables browser-to-browser
communication. PeerJS is a WebRTC client library and extensible signaling
server. Raft.js is an implementation in JavaScript of the Raft distributed
consensus protocol. These technologies are combined to create a chat
application that guarantees a shared, consistent and linearized view of the
chat history without relying on a single centralized server. We also create a
series of automated tests using the SlimerJS browser automation tool and use
them to show the correctness of the chat application and also to measure the
performance and scalability of the application over various cluster sizes.
\end{abstract}

@@

The Raft algorithm breaks the problem of consensus into three core
sub-problems: leader election, log replication, and safety. A full Raft
solution must also address: cluster membership changes, log
compaction, and client interaction. The Raft algorithm implemented and
discussed in this paper is based on "Consensus: Bridging Theory and
Practice" by Diego Ongaro (PhD Dissertation). % TODO/CITE

@@

The current Raft.js implementation is based on "Consensus", Ongaro and
implements most of the Raft algorithm including: leader election, log
replication, safety, membership changes, basic client interaction, and
read-only operation optimizations. Log compaction and full client
linearizability are not currently implemented (clients and client
transactions are not assigned unique IDs). Log compaction is important
for long-running production systems in order to limit memory growth
and allow new nodes to catch up their logs more efficiently. However
log compaction is not required for verifying overall system
correctness and does not significantly impact short-running test
configurations. Full client linearizability is important feature for
simplifying certain types of distributed applications, however the
implementation requires some significant additions to the Raft
implementation (e.g. assigning client IDs and sequence IDs and tracking
timed-out clients). For the chat application, simply adding one atomic
check-and-set state machine command is sufficient to achieve
a linearized chat history. This does impose a slight additional cost
on the client because it must retry the atomic operations until they
succeed. The author intends to implement both log compaction and
full client linearizability in the future but, however these do not
have a significant impact on the ability to test Raft combined with
WebRTC.

@@

\item Dynamic Raft.js over WebRTC: enhance the signaling server with support for multiple sessions/channels to enable separate Raft clusters with the same signaling server. Start each cluster with a single node and automatically add and remove Raft nodes as they come and go.

@@

In the Raft over WebRTC implementation each local browser is a client
of the Raft cluster. However in Raft, the leader node is the only node
that adds entries to the log so clients on follower nodes must have
a way to communicate with the leader. This is implemented as pair of
functions named clientRequest and clientRequestResponse. clientRequest
is called by the client with the arguments of a client request RPC and
a callback function to be called once the request is completed. The
clientRequest/clientRequestResponse functions work as follows:

@@

\begin{enumerate}
    \item \emph{test/common.js}: a module of common functions used by
        the other tests scripts.
    \item \emph{test/test\_up.js}: launch and start an initial leader
        node/page, then start the rest of the nodes (if any) so that
        the leader will configure them into the cluster. Then test
        then waits for every node to be recorded (applied) into the log
        of every other node of the cluster. The test reports the
        number of milliseconds for the cluster to reach this state.
    \item \emph{test/test\_kill\_nodes.js}: similar to
        \emph{test/test\_up.js} but once all the nodes are configured
        into the cluster, this script then kills the leader plus
        a configurable number of nodes. By default the script kills
        just under half the nodes of the cluster. The test then waits
        for a new leader to be elected and remove the dead nodes from
        the cluster. The test reports both the number of milliseconds
        for the cluster to come up and the recover from the failure.
    \item \emph{test/test\_chat\_propagate.js}: similar to
        \emph{test/test\_up.js} but once the cluster is up, the test
        connects to one of the non-leader nodes and and sends a chat
        text line. The test then waits for the chat message to
        propagate to every other node's state machine. The test
        reports both the number of milliseconds for the cluster to
        come up and for the messages to propagate to every node.
\end{enumerate}

@@

\begin{enumerate}
    \item \emph{CPUs}: 8 Intel Core i7 processors running at 3.20GHz
    \item \emph{Memory}: 12 GB
    \item \emph{OS}: Ubuntu 14.04.1 LTS
    \item \emph{Docker}: version 1.3.1
    \item \emph{Docker image}:
    \begin{enumerate}
        \item \emph{OS}: Ubuntu 12.04.5 LTS
        \item \emph{SlimerJS}: version 0.9.6
        \item \emph{XULRunner}: 38.0.5
    \end{enumerate}
\end{enumerate}

@@

\section*{Acknowledgments}
I would like to thank my academic supervisor, David Levine, for his
excellent input and oversight of this project. I would also like to
thank Diego Ongaro for his amazing work on the Raft protocol and paper
which made implementing a Raft implementation a pleasure. There are
too many people involved in the creation and standardization of WebRTC
but I would like to thank them for their tireless work to make
browser-to-browser communication possible. Last but not least, I would
like to thank my family for putting up with the long nights and
weekends I spent on this project. Thanks!
